Supp mat: time series paper

### 2017-07-07 writing/manuscript_draft.tex
added citations and modified code for figure 1

Long-term census data is necessary to better understand ecological
processes and make management decisions. However, census data is often
expensive, requiring a lot of people-hours and equipment. Little work
has addressed the question, when is a time series of census data long
enough to address a question of interest? Here, we explore two
approaches to address this question: through simulations and through an
empirical approach. I specifically determine the minimum time series
length required to estimate significant increases or decreases in
population abundance. Importantly, I examine the ability to detect a
trend with a set level of statistical power, which is often neglected in
ecological time series analyses. Using a simulation approach, we
demonstrate how the minimum time series length required increases with
weaker trends in abundance and with higher population variability. In
addition, we examine 868 populations of vertebrate species to determine
the minimum time required to detect changes in their abundance. We show
that 10-15 years of continuous monitoring are required in order to
achieve a high level of statistical power. Similar to
simulation-approaches, the minimum time required for field census data
strongly depends on trend strength, population variability, and temporal
autocorrelation. These results point to the importance of sampling
populations over long periods of time. Most studies, especially those
less than 10-15 years, are probably underpowered and potentially
misleading.

@@

Observational studies and population censuses have become a cornerstone
of modern ecological research and conservation biology (Magurran et al.
2010). long-term data is necessary to both understand population
dynamics and to assess extinction risk of species. Even though many time
series may now be considered ``long-term'', most are still short
(i.e.~less than X years). \emph{I could have statistic on how long is
considered ``long-term'' in papers}. Time series are typically short for
a variety of reasons. They are often coupled with an experiment, which
may only last a couple of years. In addition, short funding cycles make
it difficult to examine populations over longer periods of time
(citations). There are, of course, notable exceptions including the Isle
Royale study, Iowa corn study, cod, lynx-hair, and many others
(citations).

@@

For time series data, two general approaches can be used. First, as can
be done with experimental data, we can relax the definition of
statistical power (cite Bolker). Instead of using classic statistical
tests (e.g.~t-test), and corresponding formulas for determinig sample
size, simulation approaches can be used (Gerrodette 1987). Simple models
can be simulated with parameter values corresponding to a population of
interest. Out of a large number of simulations, power can be calculated
as the fraction of simulations that meet some criteria. The specific
criteria depend on the question at hand. Then, time series simulations
of different lengths are run to determine the minimum time series length
required to be confident that you will be able to answer your question
(cite Bolker). In addition to using simulated data, empirical time
series can also be used. We usually do not have multiple replicates of
similar populations, but we can subsample an empirical time series
(Gerber, DeMaster, and Kareiva 1999). Subsamples of different lengths
can then be evaluated to see which fraction of subsamples meet some
criteria, again a measure of statistical power. Similar to the
simulation approach, we can use these measures of power to determine the
minimum time series required for a particular question of interest.

@@

@@

We assume that each time series in our database is long enough to
include all necessary information (e.g.~variability) about the
population. In other words, each time series is a representative sample.
We first take all possible contiguous subsamples of each time series.
For example, a time series of 40 years would have 39 possible contiguous
subsamples of length 2, 38 possible contiguous subsamples of length 3,
and continuing until 1 possible contiguous subsample of length 40
(Gerber, DeMaster, and Kareiva 1999, Giron-Nava et al. (2017)). Then we
determined the \(T_{min}\) required to be confident (i.e.~have high
enough statistical power) in an estimate of the slope coefficient from
linear regression. Therefore, we run a linear regression for each
subsampled time series. Then, we determine the fraction of subsamples of
a particular length that have estimated slope coefficients which are
statistically different from zero given the long-term, or ``true``, time
series also has a significant slope. This is a measure of statistical
power (often denoted as \(1- \beta\), where \(\beta\) is the probability
of a type II error). Lastly, we determine which subsample length is
required to guarantee we have above a certain threshold of statistical
power, which we will assume is 0.8 by convention (citation).

We also examine how minimum time series calculations changed depending
on the question of interest. In the supplementary material, we show how
to make similar calculations to determine the minimum time series
required to estimate overall population growth rate (Fig.
\ref{fig:growth_rate}), the variability in population size, and future
extinction risk. Again, we took subsamples of time series data to
determine \(T_{min}\) for each of these questions.

@@

\subsection{Questions in the appendix}\label{questions-in-the-appendix}

Describe some of the results that we include in the appendix

@@

We explored two different approaches to estimate the minimum time series
length required to address a particular question of interest. We
specifically asked, what is the minimum time series length required to
determine long-term changes in abundance using linear regression? The
simulation-based approach has been suggested by others, especially in
situations more complicated than that suited for classic power analysis
(Gerrodette 1987, Johnson2015). Our simulations support past work that
longer time series are needed when the trend strength (i.e.~rate of
increase or decrease) is weak or when population variability is high
(Gerrodette 1987).

Here, we focus on a empirical approach to estimate the minimum time
series length requried to assess changes in abundance over time. We
examined long time series and then subsampled each to determine the
minimum time required to achieve a desired level of significance and
power for linear regression. Statistical power is important as it
provides on information as to the neccessary samples required to
dertermine a significant trend (Legg and Nagy 2006). We found that for
878 populations, at least 10-15 years of continuous monitoring were
neccessary (Fig. \ref{fig:min_time_dist}). This timeframe is in line
with past work on a short-lived snail species (Rueda-Cediel et al. 2015)
and a long-lived whale species (Gerber, DeMaster, and Kareiva 1999). In
line with theoretical predictions (citations), we also found \(T_{min}\)
was strongly correlated with the trend strength, variability in
population size, and temporal autocorrelation (Fig.
\ref{fig:correlates}). Against our prior expectations, we also found
that \(T_{min}\) did not correlate with any biological variables of
interest (cite figure or supp mat). We initially hypothesized that
species with longer lifespans or generation times may require a longer
sampling period. Our result could have result of at least two effects.
First, the database we used may not include a diverse enough set of
species with different life history traits. Second, the question we
poised, whether a population is increasing or decreases, is only worried
about population surveys. Therefore, biological variables may be more
important if the question itself was different. ANOTHER SENTENCE HERE

There has also been interest in determining the minimum time series
length in the bird community (Hatch 2003, ({\textbf{???}}), Seavy and
Reynolds (2007)). Hatch (2003) used seabird monitoring data to estimate
the minimum time needed to achieve appropriate levels of power. He found
that the minimum time required ranged between 11 and 69 years depending
on species, trend strength, and study design. Similarly,
({\textbf{???}}) recommended a standard of 80\% power to detect a 50\%
decline over 20 years in landbird populations. They found that only
about 42\% of populations in the Breeding Bird Survey met their
criteria. Seavy and Reynolds (2007) asked whether statistical power was
even a useful framework for assessing long-term population trends. They
used 24 years of census data on Red-tailed Tropicbirds
(\emph{Phaethon rubricauda}) in Hawaii and showed that to detect a 50\%
decline over 10 years almost always resulting in high statistical power
(above 0.8). Therefore, they cautioned against only using power analyses
to design monitoring schemes and instead argued for precision based
metrics ..

An important related question, is the optimal allocation of sampling
effort in space versus time (citations). In a theoretical investigation
of this question, Rhodes and Jonzen (2011) found that the optimal
allocation of sampling depended strongly on temporal and spatial
autocorrelation. If spatial population dynamics were highly correlated,
then it was better to sample more temporally, and vice versa. Our work
supports this idea as populations with strong temporal autocorrelation
needed less years of sampling (cite figure here). Morrison and Hik
(2008) also studied the optimal allocation of sampling effort in space
versus time, but used emprical data from a long-term census of the
collared pika (\emph{Ochotona collaris}) found in the Yukon. They
estimated long-term growth rates among three subpopulations over a
10-year period. They found that censuses less than 5 years may be
misleading and that extrapolating from one population to another, even
when nearby, may be difficult. Therefore, they argued more sampling
should be conducted spatially than temporally (Morrison and Hik 2008).

\subsection{Other questions addressed}\label{other-questions-addressed}

In the appendix we asked two additional questions to make estimates of
\(T_{min}\). First, we asked how many years of sampling are neccessary
to estimate a geometric growth rate?

@@

Our work has some limitations in determining the minimum time series
length required. First, \(T_{min}\) depends strongly on the question of
interest. For our empirical approach, the subsampling of the full time
series allows for estimates of power, but the individual subsamples are
clearly not independent of one another. In addition, estimates of
\(T_{min}\) depend on chosen values of \(\alpha\) and \(\beta\) (see
appendix). In an ideal world, we would build a specific population model
parameterized for each population of interest. Then, model simulations
could be used to estimate the minimum time series reuqired to address a
specific question of interest. Clearly, this is not always practical,
especially if running analyses for a wide array of species as we do
here. In addition, our statistical models suggest that \(T_{min}\) does
not correlate with any biological variables of interest, at least for
the question of linear regression. Therefore, it is not possible to use
these results to predict \(T_{min}\) for another population, even if the
population is of a species with a similar life-history to one in our
database.

@@

We use a database of 878 populations to determine the minimum time
series length required. We show that to identify long-term changes in
abundance, 10-15 years of continuos monitoring are required (Fig.
\ref{fig:min_time_dist}). In line with theoretical predictions, we also
show that \(T_{min}\) is strongly correlated with the overall trend in
abundance, variability in abundance, and the temporal autocorrelation
(Fig. \ref{fig:correlates}). Our work implies that for many populations,
census data less than 10-15 years are probably not reliable. This
stresses the importance of long-term monitoring programs. From both a
scientific and management perspective estimates of \(T_{min}\) are
important. If a time series is too short, we are likely to make
incorrect conclusions about long-term trends. In addition, a time series
that is too long may be a poor use of already limited funds (Gerber,
DeMaster, and Kareiva 1999). Future work can examine other species, with
a wider range of life history characteristics. In addition, similar
approaches can be used to determine the minimum time series length
required to address other questions.

@@

ERW was partially supported by a National Science Foundation Graduate
Fellowship. We would like to thank members of the Ecological Theory
group at the University of California, Davis for their insight. We would
also like to thank X anonymous reviewers for their helpful comments.

@@

\hypertarget{refs}{}
\hypertarget{ref-Gerber1999}{}
Gerber, L R, D P DeMaster, and P M Kareiva. 1999. ``Gray whales and the
value of monitering data in implementing the U.S. endangered species
act.'' \emph{Conservation Biology} 13 (5): 1215--9.

@@

@@

\hypertarget{ref-Legg2006}{}
Legg, Colin J, and Laszlo Nagy. 2006. ``Why most conservation monitoring
is , but need not be , a waste of time.'' \emph{Journal of Environmental
Management} 78: 194--99.
doi:\href{https://doi.org/10.1016/j.jenvman.2005.04.016}{10.1016/j.jenvman.2005.04.016}.

@@

\hypertarget{ref-Morrison2008}{}
Morrison, Sharn, and David S. Hik. 2008. ``When? Where? and for how
long? Census design considerations for an Alpine Lagomorpg, the Collared
pika.'' In \emph{Lagomorph Biology}, 103--13. Springer Berlin
Heidelberg.
doi:\href{https://doi.org/10.1007/978-3-540-72446-9}{10.1007/978-3-540-72446-9}.

@@

\hypertarget{ref-Rueda-Cediel2015}{}
Rueda-Cediel, Pamela, Kurt E Anderson, Tracey J Regan, Janet Franklin,
and M Regan. 2015. ``Combined Influences of Model Choice , Data Quality
, and Data Quantity When Estimating Population Trends.'' \emph{PLoSONE}
10 (7): e0132255.
doi:\href{https://doi.org/10.1371/journal.pone.0132255}{10.1371/journal.pone.0132255}.

@@

>>>>

Long-term census data is necessary to better understand ecological
processes and make management decisions. However, census data is often
expensive, requiring a lot of people-hours and equipment. Little work
has addressed the question, when is a time series of census data long
enough to address a question of interest? Here, we explore two
approaches to address this question: through simulations and through an
empirical approach. We specifically determine the minimum time series
length required to estimate significant increases or decreases in
population abundance. Importantly, We examine the ability to detect a
trend with a set level of statistical power, which is often neglected in
ecological time series analyses. Using a simulation approach, we
demonstrate how the minimum time series length required increases with
weaker trends in abundance and with higher population variability. In
addition, we examine 868 populations of vertebrate species to determine
the minimum time required to detect changes in their abundance. We show
that 10-15 years of continuous monitoring are required in order to
achieve a high level of statistical power. Similar to
simulation-approaches, the minimum time required for field census data
strongly depends on trend strength, population variability, and temporal
autocorrelation. These results point to the importance of sampling
populations over long periods of time. Most studies, especially those
less than 10-15 years, are probably underpowered and potentially
misleading.

@@

Observational studies and population censuses have become a cornerstone
of modern ecological research and conservation biology (Magurran et al.
2010; Hughes et al. 2017). Long-term data is necessary to both
understand population dynamics and to assess extinction risk of species.
Even though many time series may now be considered ``long-term'', most
are still short. Time series are typically short for a variety of
reasons. They are often coupled with an experiment, which may only last
a couple of years. In addition, short funding cycles make it difficult
to examine populations over longer periods of time (Hughes et al.
2017){]}. There are, of course, notable exceptions where long-term
monitoring has occured, for example the continuous plankton recorder
(Giron-Nava et al. 2017).

@@

For time series data, two general approaches can be used. First, as can
be done with experimental data, we can relax the definition of
statistical power (Bolker 2008). Instead of using classic statistical
tests (e.g.~t-test), and corresponding formulas for determinig sample
size, simulation approaches can be used (Gerrodette 1987, Bolker2008).
Simple models can be simulated with parameter values corresponding to a
population of interest. Out of a large number of simulations, power can
be calculated as the fraction of simulations that meet some criteria.
The specific criteria depend on the question at hand. Then, time series
simulations of different lengths are run to determine the minimum time
series length required to be confident that you will be able to answer
your question (Bolker 2008). In addition to using simulated data,
empirical time series can also be used. We usually do not have multiple
replicates of similar populations, but we can subsample an empirical
time series (Gerber, DeMaster, and Kareiva 1999). Subsamples of
different lengths can then be evaluated to see which fraction of
subsamples meet some criteria, again a measure of statistical power.
Similar to the simulation approach, we can use these measures of power
to determine the minimum time series required for a particular question
of interest.

@@

\begin{figure}[htbp]
\centering
\includegraphics{manuscript_draft_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{(a) example of simulated time series for 40 time periods and
(b) statistical power versus the simulated time series length. The
horizontal, dashed line is our desired statistical power of 0.8. The
vertical, dashed line is the minimum time required to achieve our
desired statistical power. (c) Minimum time required for different
values of the trend strength (\(r\)) and (d) minimum time required for
different levels of population variability
(\(\sigma\)).\label{fig:theoretical_approach}}
\end{figure}

@@

We assume that each time series in our database is long enough to
include all necessary information (e.g.~variability) about the
population. In other words, each time series is a representative sample.
We first take all possible contiguous subsamples of each time series.
For example, a time series of 40 years would have 39 possible contiguous
subsamples of length 2, 38 possible contiguous subsamples of length 3,
and continuing until 1 possible contiguous subsample of length 40
(Gerber, DeMaster, and Kareiva 1999; Giron-Nava et al. 2017). Then we
determined the \(T_{min}\) required to be confident (i.e.~have high
enough statistical power) in an estimate of the slope coefficient from
linear regression. Therefore, we run a linear regression for each
subsampled time series. Then, we determine the fraction of subsamples of
a particular length that have estimated slope coefficients which are
statistically different from zero given the long-term, or ``true``, time
series also has a significant slope. This is a measure of statistical
power (often denoted as \(1 - \beta\), where \(\beta\) is the
probability of a type II error). Lastly, we determine which subsample
length is required to guarantee we have above a certain threshold of
statistical power, which we will assume is 0.8 by convention (citation).

@@

\subsection{Questions in the supplmentary
material}\label{questions-in-the-supplmentary-material}

In the supplmentary material, we explore several additional questions
related to statistical power and estimating \(T_{min}\). First, we
showed that other metrics besides estimating the slope coefficient from
linear regression could be used. More specifically, we show how the same
approach desribed here can be used to determine the minimum time
required to estimate a population's geometric growth rate (Figs.
\ref{fig:growth_rate},\ref{fig:min_time_growth_dist}). We also explored
how the simulation approach can be applied to more
biologically-realistic population models (Fig. \ref{fig:shark_example}).
More specifically, we determine the minimum time required to estimate
long-term population trends using a stochastic, age-structured model of
lemon shark population dynamics in the Bahamas (White, Nagy, and Gruber
2014). We found that over 27 years of continous monitoring were needed
in this particular scenario (Fig. \ref{fig:shark_example}. Lastly, we
determine the minimum time required to estimate long-term trends
according to generalized additive models, instead of the simple linear
models used here (Fig. \ref{fig:gam_example}). We found that although we
obtain a similar distribution of minimum times required for GAMs, the
minimum time required for GAMs is on average shorter than for linear
regression (Fig. \ref{fig:min_time_dist_gam}). We expand on these
results in the supplementary material.

@@

We explored two different approaches to estimate the minimum time series
length required to address a particular question of interest. We
specifically asked, what is the minimum time series length required to
determine long-term changes in abundance using linear regression? The
simulation-based approach has been suggested by others, especially in
situations more complicated than that suited for classic power analysis
(Gerrodette 1987; P. C. Johnson et al. 2015; Bolker 2008). Our
simulations support past work that longer time series are needed when
the trend strength (i.e.~rate of increase or decrease) is weak or when
population variability is high (Gerrodette 1987).

Here, we focus on an empirical approach to estimate the minimum time
series length requried to assess changes in abundance over time. We
examined long time series and then subsampled each to determine the
minimum time required to achieve a desired level of significance and
power for linear regression. Statistical power is important as it
provides on information as to the neccessary samples required to
determine a significant trend (Legg and Nagy 2006). We found that for
878 populations, at least 10-15 years of continuous monitoring were
neccessary (Fig. \ref{fig:min_time_dist}). This timeframe is in line
with past work on a short-lived snail species (Rueda-Cediel et al. 2015)
and a long-lived whale species (Gerber, DeMaster, and Kareiva 1999). In
line with theoretical predictions (Rhodes and Jonzen 2011), we also
found \(T_{min}\) was strongly correlated with the trend strength,
variability in population size, and temporal autocorrelation (Fig.
\ref{fig:correlates}). Contrary to our prior expectations, we also found
that \(T_{min}\) did not correlate with any biological variables of
interest (Fig. \ref{fig:biological_correlates}). We initially
hypothesized that species with longer lifespans or generation times may
require a longer sampling period. Our result could have been a result of
at least two effects. First, the database we used may not include a
diverse enough set of species with different life history traits.
Second, the question we poised, whether a population is increasing or
decreasing, was specifically concerned with trends in population density
over time. Therefore, life-history characteristics may be more important
for other questions (e.g.~extinction risk) where estimates minimum time
required are needed.

There has been interest in determining the minimum time series length in
the bird research community (Hatch 2003; Bart et al. 2004; Seavy and
Reynolds 2007). For example, Hatch (2003) used seabird monitoring data
to estimate the minimum time needed to achieve appropriate levels of
power. He found that the minimum time required ranged between 11 and 69
years depending on species, trend strength, and study design. Similarly,
Bart et al. (2004) recommended a standard of 80\% power to detect a 50\%
decline over 20 years in landbird populations. They found that only
about 42\% of populations in the Breeding Bird Survey met their
criteria. Seavy and Reynolds (2007) asked whether statistical power was
even a useful framework for assessing long-term population trends. They
used 24 years of census data on Red-tailed Tropicbirds
(\emph{Phaethon rubricauda}) in Hawaii and showed that to detect a 50\%
decline over 10 years almost always resulting in high statistical power
(above 0.8). Therefore, they cautioned against only using power analyses
to design monitoring schemes and instead argued for metrics that would
increase precision. For example, Seavy and Reynolds (2007) suggest
improving randomization, reducing bias, and increases detection
probability when designing and evaluating monitoring programs.

An important related question, is the optimal allocation of sampling
effort in space versus time. In a theoretical investigation of this
question, Rhodes and Jonzen (2011) found that the optimal allocation of
sampling depended strongly on temporal and spatial autocorrelation. If
spatial population dynamics were highly correlated, then it was better
to sample more temporally, and vice versa. Our work supports this idea
as populations with strong temporal autocorrelation needed less years of
sampling (Fig. \ref{fig:correlates}). Morrison and Hik (2008) also
studied the optimal allocation of sampling effort in space versus time,
but used emprical data from a long-term census of the collared pika
(\emph{Ochotona collaris}) found in the Yukon. They estimated long-term
growth rates among three subpopulations over a 10-year period. They
found that censuses less than 5 years may be misleading and that
extrapolating from one population to another, even when nearby
geographically, may be untenable.

@@

Our work has some limitations in determining the minimum time series
length required. First, \(T_{min}\) is particular to the specific
question of interest. For example, in the supplementary material, we
used generalized additive models to detect population trends and came up
with a similar, but quantivaely different value for \(T_{min}\) (Fig.
\ref{fig:min_time_dist_gam}). An additional complication is that for our
empirical approach, the subsampling of the full time series allows for
estimates of power, but the individual subsamples are clearly not
independent of one another. Further, estimates of \(T_{min}\) depend on
chosen values of \(\alpha\) and \(\beta\) (Fig.
\ref{fig:min_time_vs_alpha_beta}). In an ideal setting, we would build a
specific population model parameterized for each population of interest.
Then, model simulations could be used to estimate the minimum time
series required to address each specific question of interest. Clearly,
this is not always practical, especially if conducting analyses for a
wide array of species as we do here. In addition, our statistical models
suggest that \(T_{min}\) does not correlate with any biological
variables of interest, at least for the question of linear regression
(Fig. \ref{fig:biological_correlates}). Therefore, it is not possible to
use these results to predict \(T_{min}\) for another population, even if
the population is of a species with a similar life-history to one in our
database.

@@

We use a database of 878 populations to determine the minimum time
series length required to determine population trends. This goes beyond
previous work that either focused on theoretical investigations or a
limited number of species. We show that to identify long-term changes in
abundance, 10-15 years of continuos monitoring are required (Fig.
\ref{fig:min_time_dist}). In line with theoretical predictions
(Gerrodette 1987), we also show that \(T_{min}\) is strongly correlated
with the long-term population trend (i.e.~rate of increase), variability
in population size, and the temporal autocorrelation (Fig.
\ref{fig:correlates}). Our work implies that for many populations,
census data less than 10-15 years are probably not reliable. This
stresses the importance of long-term monitoring programs. From both a
scientific and management perspective estimates of \(T_{min}\) are
important. If a time series is too short we lack statistical power to
reliably detect long-term population trends. In addition, a time series
that is too long may be a poor use of already limited funds (Gerber,
DeMaster, and Kareiva 1999). Future work should examine other species,
with a wider range of life history characteristics. In addition, similar
approaches can be used to determine the minimum time series length
required to address additional questions of interest.

@@

ERW was partially supported by a National Science Foundation Graduate
Fellowship. We would like to thank members of the Ecological Theory
group at the University of California, Davis for their insight. We would
also like to thank T. Dallas and E. Malcolm-Whitefor their helpful
comments.

@@

\hypertarget{refs}{}
\hypertarget{ref-Bart2004}{}
Bart, Jonathon, Kenneth P. Burnham, Erica H. Dunn, Charles M. Francis,
and C. John Ralph. 2004. ``Goals and strategies for estimating trends in
landbird abundance.'' \emph{Journal of Wildlife Management} 68 (3):
611--26.

\hypertarget{ref-Bolker2008}{}
Bolker, Benjamin M. 2008. \emph{Ecological Models and Data in R}. 1st
ed. Princeton, New Jersey: Princeton University Press.

\hypertarget{ref-Gerber1999}{}
Gerber, L R, D P DeMaster, and P M Kareiva. 1999. ``Gray whales and the
value of monitering data in implementing the U.S. endangered species
act.'' \emph{Conservation Biology} 13 (5): 1215--9.

@@

\hypertarget{ref-Hughes2017}{}
Hughes, Brent B, Rodrigo Beas-luna, Allison K Barner, Kimberly Brewitt,
Daniel R Brumbaugh, Elizabeth B. Cerny-Chipman, Sarah L. Close, et al.
2017. ``Long-term studies contribute disproportionately to ecology and
policy.'' \emph{BioScience} 67 (3): 271--81.
doi:\href{https://doi.org/10.1093/biosci/biw185}{10.1093/biosci/biw185}.

@@

\hypertarget{ref-Legg2006}{}
Legg, Colin J, and Laszlo Nagy. 2006. ``Why most conservation monitoring
is, but need not be, a waste of time.'' \emph{Journal of Environmental
Management} 78: 194--99.
doi:\href{https://doi.org/10.1016/j.jenvman.2005.04.016}{10.1016/j.jenvman.2005.04.016}.

@@

\hypertarget{ref-Morrison2008}{}
Morrison, Sharn, and David S. Hik. 2008. ``When? Where? And for how
long? Census design considerations for an Alpine Lagomorph, the Collared
pika.'' In \emph{Lagomorph Biology}, 103--13. Springer Berlin
Heidelberg.
doi:\href{https://doi.org/10.1007/978-3-540-72446-9}{10.1007/978-3-540-72446-9}.

@@

\hypertarget{ref-Rueda-Cediel2015}{}
Rueda-Cediel, Pamela, Kurt E Anderson, Tracey J Regan, Janet Franklin,
and M Regan. 2015. ``Combined influences of model choice, data quality,
and data quantity when estimating population trends.'' \emph{PLoSONE} 10
(7): e0132255.
doi:\href{https://doi.org/10.1371/journal.pone.0132255}{10.1371/journal.pone.0132255}.

@@

\hypertarget{ref-White2014}{}
White, Easton R, John D Nagy, and Samuel H Gruber. 2014. ``Modeling the
population dynamics of lemon sharks.'' \emph{Biology Direct} 9 (23):
1--18.
